---
title: "Bingyu"
author: "Bingyu Sun"
date: "12/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

##Data Import

```{r}
cancer_raw =
  read_csv("./data/Cancer_Registry.csv") %>% 
  janitor::clean_names() %>% 
  select(target_death_rate, geography, everything()) %>% 
  separate(geography, into = c("county", "state"), sep = ",")
```

**Variables**

* **target_death_rate:** mean per capita (100,000) cancer mortalities (a)
* **avg_ann_count:** mean number of reported cases of cancer diagnosed annually (a)
* **avg_deaths_per_year:** mean number of reported mortalities due to cancer (a)
* **incidence_rate:** mean per capita (100,000) cancer diagnoses (a)
* **med_income:** median income per county (b)
* **pop_est2015:** population of county (b)
* **poverty_percent:** percent of population in poverty (b)
* **study_per_cap** per capita number of cancer-related clinical trials per county (a)
* **binned_inc:** median income per capita binned by decile (b)
* **median_age:** median age of county residents (b)
* **median_age_male:** median age of male county residents (b)
* **median_age_female:** median age of female county residents (b)
* **geography:** county name (b)
* **avg_household_size:** mean household size of county (b)
* **percent_married:** percent of county residents who are married (b)
* **pct_no_hs18_24:** percent of county residents ages 18-24 highest education attained: less than high school (b) 
* **pct_hs18_24:** percent of county residents ages 18-24 highest education attained: high school diploma (b)
* **pct_some_col18_24:** percent of county residents ages 18-24 highest education attained: some college (b) 
* **pct_bach_deg18_24:** percent of county residents ages 18-24 highest education attained: bachelor's degree (b)
* **pct_hs25_over:** percent of county residents ages 25 and over highest education attained: high school diploma (b)              
* **pct_bach_deg25_over:** percent of county residents ages 25 and over highest education attained: bachelor's degree (b)
* **pct_employed16_over:** percent of county residents ages 16 and over employed (b)        
* **pct_unemployed16_over:** percent of county residents ages 16 and over unemployed (b)     
* **pct_private_coverage:** percent of county residents with private health coverage (b)
* **pct_private_coverage_alone:** percent of county residents with private health coverage alone (no public assistance) (b) 
* **pct_emp_priv_coverage:** percent of county residents with employee-provided private health coverage (b)     
* **pct_public_coverage:** percent of county residents with government-provided health coverage (b)
* **pct_public_coverage_alone:** percent of county residents with government-provided health coverage alone (b)  
* **pct_white:** percent of county residents who identify as White (b)           
* **pct_black:** percent of county residents who identify as Black (b)
* **pct_asian:** percent of county residents who identify as Asian (b)                  
* **pct_other_race:** percent of county residents who identify in a category which is not White, Black, or Asian (b)          
* **pct_married_households:** percent of married households (b)
* **birth_rate:** number of live births relative to number of women in county (b)

##Check raw data

```{r}
str(cancer_raw)
```

##Cancer by state
```{r}
#cancer_state_all =
  cancer_raw %>% 
  #select(-county) %>% 
  group_by(state) %>% 
  ggplot(aes(x = state, y = target_death_rate)) +
  geom_boxplot()
```

