---
title: "Eleanor_summary"
author: "Eleanor Zhang"
date: "12/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Read Data

Read original data

```{r}
cancer_data <- read_csv("./data/Cancer_Registry.csv")
summary(cancer_data)
str(cancer_data)
# check strings entry
anyNA(cancer_data$binnedInc)
anyNA(cancer_data$Geography)
View(cancer_data)
```


About this data:

*   NA variables: PctSomeCol18_24(2285/), PctEmployed16_Over(152/), PctPrivateCoverageAlone(609)
*   3047 observations, 34 variables, we selected 20 variables (after separating county and state name)
*   >25 aged information are not intact so we will not use those data

Look at the distribution of all variables (histogram)

```{r}
cancer_data %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

Select variables

```{r select variables, eval=FALSE}
cancer <- cancer_data %>% 
  janitor::clean_names() %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>% # 20 variables
  select(target_death_rate, # response: Mean per capita (100000) cancer mortalities
         med_income, # median income per county
         poverty_percent, # percent of population in poverty
         median_age, # median age of county residents
         median_age_male, median_age_female, # of county residents
         county, state, 
         pct_no_hs18_24, pct_hs18_24, # age 18-24, percent of highest diploma less than/is high school 
         pct_some_col18_24, pct_bach_deg18_24, # percent of highest diploma:some colledge/bachelor
         # pct_hs25_over, pct_bach_deg25_over, # >age 25, highest diploma: high school/bachelor
         pct_unemployed16_over, #pct_public_coverage, 
         pct_married_households,
         pct_white, pct_black, pct_asian, pct_other_race,
         incidence_rate) %>%  # percent of race
  mutate(pct_upto_hs18_24 = pct_no_hs18_24 + pct_hs18_24,
         pct_above_hs18_24 = 100 - pct_upto_hs18_24) %>% 
  select(-pct_no_hs18_24, -pct_hs18_24, -pct_some_col18_24, -pct_bach_deg18_24)
```

Select variables

```{r}
cancer_county <- cancer_data %>% 
  janitor::clean_names() %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>% 
  select(target_death_rate, incidence_rate, med_income, poverty_percent, median_age:median_age_female, avg_household_size:pct_married_households) %>% 
  select(-pct_hs25_over, -pct_bach_deg25_over, -pct_employed16_over, -percent_married) %>% 
  mutate(pct_upto_hs18_24 = pct_no_hs18_24 + pct_hs18_24,
         pct_above_hs18_24 = 100 - pct_upto_hs18_24,
         pct_with_coverage = pct_private_coverage + pct_public_coverage_alone,
         income_level = ifelse(med_income < 35000, 0, 1)) %>% 
  select(-(pct_no_hs18_24:pct_bach_deg18_24), -pct_above_hs18_24, -(pct_private_coverage:pct_public_coverage_alone)) %>% 
  na.omit

summary(cancer_county)
str(cancer_county)
View(cancer_county)
```

check NA

```{r}
anyNA(cancer_county) # No NA involved
```


Look at distribution of selected variables:

```{r variables distribution}
cancer_county %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```


plot pairs of variables

```{r}
pairs(cancer_county)
```


## Model selection (with income levels)

### data variables description

create dataset for modeling

```{r}
cancer_county_model <- cancer_county %>% select(-med_income)
names(cancer_county_model)
```

 [1] "target_death_rate"      
 [2] "incidence_rate"         
 [3] "poverty_percent"        
 [4]  "median_age"             
 [5]  "median_age_male"       
 [6] "median_age_female"      
 [7]  "avg_household_size"     
 [8]  "pct_unemployed16_over" 
[9]  "pct_white"              
[10]  "pct_black"              
[11]  "pct_asian"             
[12] "pct_other_race"         
[13]  "pct_married_households" 
[14]  "pct_upto_hs18_24"      
[15] "pct_with_coverage"     
[16]  "income_level"


look at variables

```{r}
pairs(cancer_county_model)
```

all variables

```{r}
all_var <- lm(target_death_rate ~. , data = cancer_county_model)
summary(all_var)
HH::vif(all_var)
```

backward selection

```{r}
step(all_var, direction='both')
step(all_var, direction='backward')
```

backward model

```{r}
back_model <- lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_level, data = cancer_county_model)
summary(back_model)
plot(back_model)
HH::vif(back_model)
MASS::boxcox(back_model)
```



```{r}
cancer_county_model %>% 
  group_by(income_level) %>% 
  ggplot(aes(x = poverty_percent, y = target_death_rate)) + geom_point() + geom_smooth() + facet_grid(~income_level)
```

interaction : income_level vs. poverty_percent

```{r}
lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_level + income_level*poverty_percent, data = cancer_county_model) %>% 
  summary()
```

separate by income level

```{r}
income_low_data <- cancer_county_model %>% filter(income_level == 0)
income_high_data <- cancer_county_model %>% filter(income_level == 1)

model_low_income <- lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage, data = income_low_data)

model_high_income <- lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage, data = income_high_data)

summary_table = cbind(summary(model_low_income)$coefficients[,1],
                      summary(model_low_income)$coefficients[,4],
                      summary(model_high_income)$coefficients[,1], 
                      summary(model_high_income)$coefficients[,4])
summary_table

plot(model_high_income)
plot(model_low_income)
dim(income_high_data)
dim(income_low_data)
```

## Cross validation

```{r}
library(boot)

glm.fit <- glm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_level, data = cancer_county_model)

cv.err <- cv.glm(cancer_county_model, glm.fit)

cv.err$delta
summary(back_model)
anova(back_model)
dim(cancer_county_model)
glance(back_model)
```



