---
title: "Eleanor_summary"
author: "Eleanor Zhang"
date: "12/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Read Data

Read original data

```{r}
cancer_data <- read_csv("./data/Cancer_Registry.csv")
summary(cancer_data)
str(cancer_data)
# check strings entry
anyNA(cancer_data$binnedInc)
anyNA(cancer_data$Geography)
View(cancer_data)
```


About this data:

*   NA variables: PctSomeCol18_24(2285/), PctEmployed16_Over(152/), PctPrivateCoverageAlone(609)
*   3047 observations, 34 variables, we selected 20 variables (after separating county and state name)
*   18-24 ages information are not intact so we will not use those data

Look at the distribution of all variables (histogram)

```{r}
cancer_data %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

Select variables

```{r select variables}
cancer <- cancer_data %>% 
  janitor::clean_names() %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>% # 20 variables
  select(target_death_rate, # response: Mean per capita (100000) cancer mortalities
         med_income, # median income per county
         poverty_percent, # percent of population in poverty
         median_age, # median age of county residents
         median_age_male, median_age_female, # of county residents
         county, state, 
         pct_no_hs18_24, pct_hs18_24, # age 18-24, percent of highest diploma less than/is high school 
         pct_some_col18_24, pct_bach_deg18_24, # percent of highest diploma:some colledge/bachelor
         # pct_hs25_over, pct_bach_deg25_over, # >age 25, highest diploma: high school/bachelor
         pct_unemployed16_over, #pct_public_coverage, 
         pct_married_households,
         pct_white, pct_black, pct_asian, pct_other_race,
         incidence_rate) %>%  # percent of race
  mutate(pct_upto_hs18_24 = pct_no_hs18_24 + pct_hs18_24,
         pct_above_hs18_24 = 100 - pct_upto_hs18_24) %>% 
  select(-pct_no_hs18_24, -pct_hs18_24, -pct_some_col18_24, -pct_bach_deg18_24)
summary(cancer)
str(cancer)
View(cancer)
```

check NA

```{r}
anyNA(cancer) # No NA involved
str(cancer)
```


Look at distribution of selected variables:

```{r variables distribution}
cancer %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

Income level might need adjusted!!

plot pairs of variables

```{r correlation of variables}
pairs(cancer[-c(7,8)]) # remove strings county and state
abs(cor(cancer[-c(7,8)])) > 0.6 # colinearity 
```

## Model selection (with income unchanged)

create dataset for modeling

```{r}
cancer_county <- cancer %>% select(-county, -state, -pct_above_hs18_24, -median_age)
cancer_county
```

look at variables

```{r}
pairs(cancer_county)
```

all variables

```{r}
all_var <- lm(target_death_rate ~. , data = cancer_county)
summary(all_var)
HH::vif(all_var)
```

backward selection

```{r}
step(all_var, direction='both')
step(all_var, direction='backward')
```

backward model

```{r}
back_model <- lm(target_death_rate ~ med_income + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_other_race + 
    incidence_rate + pct_upto_hs18_24, data = cancer_county)
summary(back_model)
plot(back_model)
HH::vif(back_model)
MASS::boxcox(back_model)
```


## Model Selection (with income levels)


