---
title: "Final Project"
author: "Justin Hsie, Bingyu Sun, Eleanor Zhang, Annie Yu"
date: "12/15/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(multcomp)
library(faraway)
library(dplyr)
library(patchwork)
library("leaps")
library(caret)
library(glmnet)
library(patchwork)
library(broom)
library(modelr)
library(olsrr)
library(boot)
```

##Data Import
```{r, message=FALSE}
cancer_raw =
  read_csv("./data/Cancer_Registry.csv") %>% 
  janitor::clean_names() %>% 
  dplyr::select(target_death_rate, geography, everything()) %>% 
  separate(geography, into = c("county", "state"), sep = ",")
```

## Data varibale dictionary:

* **target_death_rate:** mean per capita (100,000) cancer mortalities (a)
* **avg_ann_count:** mean number of reported cases of cancer diagnosed annually (a)
* **avg_deaths_per_year:** mean number of reported mortalities due to cancer (a)
* **incidence_rate:** mean per capita (100,000) cancer diagnoses (a)
* **med_income:** median income per county (b)
* **pop_est2015:** population of county (b)
* **poverty_percent:** percent of population in poverty (b)
* **study_per_cap** per capita number of cancer-related clinical trials per county (a)
* **binned_inc:** median income per capita binned by decile (b)
* **median_age:** median age of county residents (b)
* **median_age_male:** median age of male county residents (b)
* **median_age_female:** median age of female county residents (b)
* **geography:** county name (b)
* **avg_household_size:** mean household size of county (b)
* **percent_married:** percent of county residents who are married (b)
* **pct_no_hs18_24:** percent of county residents ages 18-24 highest education attained: less than high school (b) 
* **pct_hs18_24:** percent of county residents ages 18-24 highest education attained: high school diploma (b)
* **pct_some_col18_24:** percent of county residents ages 18-24 highest education attained: some college (b) 
* **pct_bach_deg18_24:** percent of county residents ages 18-24 highest education attained: bachelor's degree (b)
* **pct_hs25_over:** percent of county residents ages 25 and over highest education attained: high school diploma (b)              
* **pct_bach_deg25_over:** percent of county residents ages 25 and over highest education attained: bachelor's degree (b)
* **pct_employed16_over:** percent of county residents ages 16 and over employed (b)        
* **pct_unemployed16_over:** percent of county residents ages 16 and over unemployed (b)     
* **pct_private_coverage:** percent of county residents with private health coverage (b)
* **pct_private_coverage_alone:** percent of county residents with private health coverage alone (no public assistance) (b) 
* **pct_emp_priv_coverage:** percent of county residents with employee-provided private health coverage (b)     
* **pct_public_coverage:** percent of county residents with government-provided health coverage (b)
* **pct_public_coverage_alone:** percent of county residents with government-provided health coverage alone (b)  
* **pct_white:** percent of county residents who identify as White (b)           
* **pct_black:** percent of county residents who identify as Black (b)
* **pct_asian:** percent of county residents who identify as Asian (b)                  
* **pct_other_race:** percent of county residents who identify in a category which is not White, Black, or Asian (b)          
* **pct_married_households:** percent of married households (b)
* **birth_rate:** number of live births relative to number of women in county (b)

## Look at the distribution of all varibales:
```{r}
cancer_raw %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins = 30)
```

## Choose variables:
```{r}
cancer_county =
  cancer_raw %>%
  janitor::clean_names() %>% 
  dplyr::select(target_death_rate, incidence_rate, med_income, poverty_percent, median_age:median_age_female, avg_household_size:pct_married_households) %>% 
  dplyr::select(-pct_hs25_over, -pct_bach_deg25_over, -pct_employed16_over, -percent_married) %>% 
  mutate(pct_upto_hs18_24 = pct_no_hs18_24 + pct_hs18_24,
         pct_above_hs18_24 = 100 - pct_upto_hs18_24,
         pct_with_coverage = pct_private_coverage + pct_public_coverage_alone,
         income_cat = ifelse(med_income < 35000, 0, 1)) %>% 
  dplyr::select(-(pct_no_hs18_24:pct_bach_deg18_24), -pct_above_hs18_24, -(pct_private_coverage:pct_public_coverage_alone), -med_income) %>% 
  na.omit
```

## Check correlation and distribution:
```{r}
cor(cancer_county) %>% 
  knitr::kable()

cancer_county %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins = 30)
```

##The discriptive statistics:

```{r}
state_summary = function(x){
mean = mean(x)
max = max(x)
min = min(x)
median = median(x)
var = var(x)
sd = sd(x) 
sample_size=length(x)-sum(is.na(x))
tibble(mean, max, min, median, var, sd, sample_size) 
} 

df_target_death_rate <-state_summary(cancer_county$target_death_rate)
df_incidence_rate <-state_summary(cancer_county$incidence_rate)
df_poverty_percent<-state_summary(cancer_county$poverty_percent) 
df_median_age<-state_summary(cancer_county$median_age) 
df_median_agemale<-state_summary(cancer_county$median_age_male) 
df_median_agefemale<-state_summary(cancer_county$median_age_female)
df_avg_household_size<-state_summary(cancer_county$avg_household_size)
df_pct_unemployed16_over<-state_summary(cancer_county$pct_unemployed16_over) 
df_pct_white<-state_summary(cancer_county$pct_white)
df_pct_black<-state_summary(cancer_county$pct_black) 
df_pct_asian<-state_summary(cancer_county$pct_asian)
df_pct_other_race<-state_summary(cancer_county$pct_other_race)
df_pct_married_households<-state_summary(cancer_county$pct_married_households)
df_pct_upto_hs18_24<-state_summary(cancer_county$pct_upto_hs18_24) 
df_pct_with_coverage<-state_summary(cancer_county$pct_with_coverage)

state_des <- bind_rows(df_target_death_rate,
                       df_incidence_rate,
                       df_poverty_percent, 
                       df_median_age,
                       df_median_agemale,
                       df_median_agefemale,
                       df_avg_household_size,
                       df_pct_unemployed16_over, 
                       df_pct_white,
                       df_pct_black, 
                       df_pct_asian,
                       df_pct_other_race,
                       df_pct_married_households,
                       df_pct_upto_hs18_24,
                       df_pct_with_coverage) 
variable<- c("target_death_rate", "incidence_rate","poverty_percent", "median_age", "median_agemale", "median_agefemale", "avg_household_size","pct_unemployed16_over", "pct_white", "pct_black", "pct_asian", "pct_other_race", "pct_married_households", "pct_upto_hs18_24", "pct_with_coverage")

state_wholedes <- cbind(variable, state_des) 

knitr::kable(state_wholedes)
```


##Model building:

####Stepwise:
```{r}
# building full model
full_model <- lm(target_death_rate ~., data = cancer_county)
summary(full_model)

# Using the stepwise 
stepwise_model = stepAIC(full_model, direction = "both", trace = FALSE)
summary(stepwise_model)

vif(stepwise_model) 

# Cp and AIC and Adjusted R2

model_dig <- glance(stepwise_model) %>% 
  as.data.frame() %>% 
  dplyr::select(adj.r.squared, sigma, p.value, AIC, BIC) %>% 
  rename(RES = sigma) %>% 
  mutate(cp = ols_mallows_cp(stepwise_model, full_model))

model_dig
```

#### forword elimination:
```{r}
step_null <- lm(target_death_rate ~1, data = cancer_county)
for_model <- step(step_null, scope = list(lower = step_null, upper = full_model), direction = "forward")

summary(for_model)
vif(for_model)
```

##### backward elimination:
```{r}
back_model <- step(full_model, direction = "backward", trace = FALSE)
summary(back_model) # same as stepwise
vif(back_model)
```


## Cross validation and Criterion method summery
```{r}
cross_df <- crossv_mc(cancer_county, n = 100, test = 0.2)

cross_result <- cross_df %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble),
         step_mod = map(train, ~lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_cat, data = .x)),
    
    rmse_train = map2_dbl(step_mod, train, ~rmse(model = .x, data = .y)),
    rmse_test = map2_dbl(step_mod, test, ~rmse(model = .x, data = .y)))
  

mse_result = cross_result %>% 
  dplyr::select(rmse_train, rmse_test) %>%
  summarize(mse_train = (mean(rmse_train))^2,
            mse_test = (mean(rmse_test))^2) %>% 
  knitr::kable(digit = 3)
  
mse_result 

#CV
glm.fit = glm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_cat, data = cancer_county)


library(boot)                          # For cv.glm()
cv.err = cv.glm(cancer_county, glm.fit)

# The two delta values should be similar: we use the first one
# The second value is bias corrected
cv.err$delta   #434.8355

anova(back_model)[12, 3] #MSE: 431
```

##Bootstrap
```{r}
back_model = lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_cat, data = cancer_county)

cancer_county %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(models = map(strap, ~lm(target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_male + pct_unemployed16_over + pct_white + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage + income_cat, data = .x)),
         results = map(models, broom::tidy)) %>% 
  dplyr::select(-strap, -models) %>% 
  unnest() %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))

```



## Outliers
```{r, message=FALSE}
par(mfrow = c(2,2))
plot(stepwise_model) # we observe the 282, 1059 are outliers

# Using the studentized residuals
stu_res_step <- rstandard(stepwise_model)
outliers_y_step = stu_res_step[abs(stu_res_step) > 2.5]
outliers_y_step

stu_res_for = rstandard(for_model)
outliers_y_for = stu_res_for[abs(stu_res_for) > 2.5]
outliers_y_for

# various measure of influence
influence.measures(stepwise_model) #we observe the 282, 1059 are potential outliers
```

# Remove outliers:
```{r}

# filter outlier 282
step_model_no_282 <- cancer_county %>% 
  dplyr::select(-median_age, -pct_black, -median_age_female, avg_household_size) %>% 
  tibble::rowid_to_column() %>% 
  filter(rowid !=282) %>% 
  na.omit()

step_no_282 <- update(stepwise_model, . ~ ., data = step_model_no_282)
summary(step_no_282)

# filter #282, 1059
step_model_no_all <- cancer_county %>% 
  dplyr::select(-median_age, -pct_black, -median_age_female, avg_household_size) %>% 
  tibble::rowid_to_column() %>% 
  filter(rowid !=282 & rowid != 1059) %>% 
  na.omit()

step_no_all <- update(stepwise_model, . ~ ., data = step_model_no_all)
summary(step_no_all)

par(mfrow = c(2,2))
plot(step_no_all)
```

