---
title: "model selection"
author: "Eleanor Zhang"
date: "12/16/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(leaps)
library(caret)
library(glmnet)
library(broom)
library(modelr)
library(boot)
```

## read data and select variables to import

```{r}
cancer_data <- read_csv("./data/Cancer_Registry.csv")

cancer_county <- cancer_data %>% 
  janitor::clean_names() %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>% 
  dplyr::select(target_death_rate, incidence_rate, med_income, poverty_percent, median_age:median_age_female, avg_household_size:pct_married_households) %>% 
  dplyr::select(-pct_hs25_over, -pct_bach_deg25_over, -pct_employed16_over, -percent_married) %>% 
  mutate(pct_upto_hs18_24 = pct_no_hs18_24 + pct_hs18_24,
         pct_above_hs18_24 = 100 - pct_upto_hs18_24,
         pct_with_coverage = pct_private_coverage + pct_public_coverage_alone,
         income_cat = ifelse(med_income < 35000, 0, 1)) %>% 
  dplyr::select(-(pct_no_hs18_24:pct_bach_deg18_24), -pct_above_hs18_24, -(pct_private_coverage:pct_public_coverage_alone)) %>% 
  na.omit

dim(cancer_county)
anyNA(cancer_county)
```


## Data description:

* **target_death_rate:** mean per capita (100,000) cancer mortalities (a)
* **avg_ann_count:** mean number of reported cases of cancer diagnosed annually (a)
* **avg_deaths_per_year:** mean number of reported mortalities due to cancer (a)
* **incidence_rate:** mean per capita (100,000) cancer diagnoses (a)
* **med_income:** median income per county (b)
* **pop_est2015:** population of county (b)
* **poverty_percent:** percent of population in poverty (b)
* **study_per_cap** per capita number of cancer-related clinical trials per county (a)
* **binned_inc:** median income per capita binned by decile (b)
* **median_age:** median age of county residents (b)
* **median_age_male:** median age of male county residents (b)
* **median_age_female:** median age of female county residents (b)
* **geography:** county name (b)
* **avg_household_size:** mean household size of county (b)
* **percent_married:** percent of county residents who are married (b)
* **pct_no_hs18_24:** percent of county residents ages 18-24 highest education attained: less than high school (b) 
* **pct_hs18_24:** percent of county residents ages 18-24 highest education attained: high school diploma (b)
* **pct_some_col18_24:** percent of county residents ages 18-24 highest education attained: some college (b) 
* **pct_bach_deg18_24:** percent of county residents ages 18-24 highest education attained: bachelor's degree (b)
* **pct_hs25_over:** percent of county residents ages 25 and over highest education attained: high school diploma (b)              
* **pct_bach_deg25_over:** percent of county residents ages 25 and over highest education attained: bachelor's degree (b)
* **pct_employed16_over:** percent of county residents ages 16 and over employed (b)        
* **pct_unemployed16_over:** percent of county residents ages 16 and over unemployed (b)     
* **pct_private_coverage:** percent of county residents with private health coverage (b)
* **pct_private_coverage_alone:** percent of county residents with private health coverage alone (no public assistance) (b) 
* **pct_emp_priv_coverage:** percent of county residents with employee-provided private health coverage (b)     
* **pct_public_coverage:** percent of county residents with government-provided health coverage (b)
* **pct_public_coverage_alone:** percent of county residents with government-provided health coverage alone (b)  
* **pct_white:** percent of county residents who identify as White (b)           
* **pct_black:** percent of county residents who identify as Black (b)
* **pct_asian:** percent of county residents who identify as Asian (b)                  
* **pct_other_race:** percent of county residents who identify in a category which is not White, Black, or Asian (b)          
* **pct_married_households:** percent of married households (b)
* **birth_rate:** number of live births relative to number of women in county (b)

Separate entire dataset into two groups: low income and high income:

```{r}
income_low_data <- cancer_county %>% filter(income_cat == 0) %>% dplyr::select(-income_cat)
income_high_data <- cancer_county %>% filter(income_cat == 1) %>% dplyr::select(-income_cat)
```

Description

```{r}
summary(income_low_data)
summary(income_high_data)

income_low_data %>%
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins = 30)

income_high_data %>%
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins = 30)

cor(income_low_data) %>% 
  knitr::kable()

cor(income_high_data) %>% 
  knitr::kable()
```

## Model selection

#### full model

```{r}
# low income
full_model_low <- lm(target_death_rate ~., data = income_low_data)
summary(full_model_low)

# high income
full_model_high <- lm(target_death_rate ~., data = income_high_data)
summary(full_model_high)
```


#### step wise procedure

```{r}
# low income
step(full_model_low, direction = "backward")

backward_model_low <- lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + 
    pct_asian + pct_other_race + pct_upto_hs18_24 + pct_with_coverage, 
    data = income_low_data)
summary(backward_model_low)

# high income
step(full_model_high, direction = "backward")

backward_model_high <- lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + pct_unemployed16_over + pct_black + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage, data = income_high_data)
summary(backward_model_high)
```

Check assumption

```{r}
plot(backward_model_low)
plot(backward_model_high)
```

#### citerion based approach

low income
```{r}
library(leaps)
criterion_stats_low =
  regsubsets(target_death_rate ~ ., nvmax = 12, data = income_low_data) %>% 
  summary()

criterion_stats_low

par(mar = c(4,4,1,1))
par(mfrow = c(1,2))

plot(2:13, criterion_stats_low$cp, xlab = "No of parameters", ylab = "Cp Statistic")
abline(0,1)

plot(2:13, criterion_stats_low$adjr2, xlab = "No of parameters", ylab = "Adj R2")

## 9 predictors are sufficient
criterion_model_low <- lm(formula = target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = income_low_data)
```

high income

```{r}
criterion_stats_high =
  regsubsets(target_death_rate ~ ., nvmax = 12, data = income_high_data) %>% 
  summary()

criterion_stats_high
par(mar = c(4,4,1,1))
par(mfrow = c(1,2))

plot(2:13, criterion_stats_high$cp, xlab = "No of parameters", ylab = "Cp Statistic")
abline(0,1)

plot(2:13, criterion_stats_high$adjr2, xlab = "No of parameters", ylab = "Adj R2")

# 7 predictors are sufficients
criterion_model_high <- lm(formula = target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, 
    data = income_high_data)
```

check assumption and influential points

```{r}
plot(criterion_model_low)
plot(criterion_model_high)
```


## Model comparison

backward_model_low <- lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + 
    pct_asian + pct_other_race + pct_upto_hs18_24 + pct_with_coverage, 
    data = income_low_data)  10 predictor
    
criterion_model_low <- lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = income_low_data)  9 predictor
    

backward_model_high <- lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + pct_unemployed16_over + pct_black + pct_asian + 
    pct_other_race + pct_married_households + pct_upto_hs18_24 + 
    pct_with_coverage, data = income_high_data) 10 predictor


criterion_model_high <- lm(target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, 
    data = income_high_data)  7 predictor

#### select model with least predictors

```{r}
low_model <- lm(formula = target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = income_low_data)
high_model <- lm(formula = target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, 
    data = income_high_data)
plot(low_model)
plot(high_model)
```

## Influential points

remove influential points in low income
```{r}
income_low_rm <- income_low_data[-c(30,101,332),]

low_model_rm<- lm(formula = target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = income_low_rm)
summary(low_model_rm)
summary(low_model)
```

remove influential points in high income

```{r}
income_high_rm <- income_high_data[-c(1088, 1204, 248),]

high_model_rm <- lm(formula = target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, 
    data = income_high_rm)
summary(high_model_rm)
summary(high_model)
```


## cross validation

CV for low and high income model

```{r}
cross_df_low = crossv_mc(income_low_data, n = 100, test = 0.2) 
cross_df_high = crossv_mc(income_high_data, n = 100, test = 0.2) 

cross_result_low = 
  cross_df_low %>% 
  mutate(
    step_mod = map(train, ~lm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = .x)),
    rmse_train = map2_dbl(step_mod, train, ~rmse(model = .x, data = .y)),
    rmse_test = map2_dbl(step_mod, test, ~rmse(model = .x, data = .y))
    )
  

mse_results_low = cross_result_low %>% 
  dplyr::select(rmse_train, rmse_test) %>%
  summarize(mse_train_low = (mean(rmse_train))^2,
            mse_test_low = (mean(rmse_test))^2) #mse results

cross_result_high = 
  cross_df_high %>% 
  mutate(
    step_mod = map(train, ~lm(target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, data = .x)),
    rmse_train = map2_dbl(step_mod, train, ~rmse(model = .x, data = .y)),
    rmse_test = map2_dbl(step_mod, test, ~rmse(model = .x, data = .y))
    )
  

mse_results_high = cross_result_high %>% 
  dplyr::select(rmse_train, rmse_test) %>%
  summarize(mse_train_high = (mean(rmse_train))^2,
            mse_test_high = (mean(rmse_test))^2)

#LOOCV
glm.fit_low = glm(target_death_rate ~ incidence_rate + med_income + 
    median_age_male + median_age_female + pct_white + pct_black + pct_asian + 
    pct_other_race +  pct_with_coverage, data = income_low_data)

cv.err_low = cv.glm(income_low_data, glm.fit_low)

glm.fit_high = glm(target_death_rate ~ incidence_rate + med_income +
                             median_age_female + pct_unemployed16_over + pct_black + pct_other_race +
                             pct_upto_hs18_24, data = income_high_data)

cv.err_high = cv.glm(income_high_data, glm.fit_high)

# The two delta values should be similar: we use the first one
# The second value is bias corrected
cv.err_low$delta  

anova(low_model)
anova(low_model)[10, 3] #MSE: 431
anova(high_model)[8, 3]

mse_low = 
  tibble(
  mse_model = anova(low_model)[10, 3],
  mse_LOOCV = cv.err_low$delta[1],
  mse_CV_train = mse_results_low$mse_train_low,
  mse_CV_test = mse_results_low$mse_test_low
)

mse_high = 
  tibble(
  mse_model = anova(high_model)[8, 3],
  mse_LOOCV = cv.err_high$delta[1],
  mse_CV_train = mse_results_high$mse_train_high,
  mse_CV_test = mse_results_high$mse_test_high
)

rbind(mse_low, mse_high) %>% mutate(dataset = c("low income", "high income")) %>% 
  dplyr::select(dataset, everything()) %>% knitr::kable(digits = 3)
```



